Respiratory-Related Mortality Rates Show A Positive Correlation With Increasing Air Pollution

### 2024-03-16 paper/paper.qmd
FINAL PAPER

---
title: "Respiratory-Related Mortality Rates Show A Positive Correlation With Increasing Air Pollution"
subtitle: "Based on Data Collected From The Province Of Alberta"
author: 
  - Vanshika Vanshika
  - Shivank Goel
  - Navya Hooda
thanks: "Code and data are available at: https://github.com/shivankgoel003/Mortality-in-Alberta."
date: today
date-format: long
abstract: "Air pollution's severe impact on human respiratory and cardiac health is a growing public health concern across the world. The fine particulate matter (PM2.5) air pollutant is of particular concern because it can deeply penetrate the lungs and cause systemic effects and is largely found in urban environments. This research uses data from Alberta, Canada to assess any correlations between PM2.5 levels and the regional Air Quality Health Index (AQHI), against the number of deaths from groups of respiratory and cardiac illnesses we selected. We utilized statistical models and identified a strong association between increased PM2.5 levels and higher mortality from these diseases. These findings suggest that there is an immediate need for health policies that address air quality improvements to safeguard public health in Alberta and beyond."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

In our study, there might be several limitations affecting our conclusions.
First, between 2010 and 2017, older equipment at Alberta’s monitoring stations was replaced with new monitoring equipment for PM2.5. These new instruments measured an additional portion (semi-volatile) of the PM2.5 mass not captured by older instruments. As a result, the older monitoring equipment used between 2000 and 2017 likely underreported concentrations of PM2.5 under some conditions, after 2010, the increase in PM2.5 concentrations may be a result of changes in monitoring equipment (PM2.5 reference). Overall, the concentrations measured with the new monitors may not be directly comparable with measurements from years in which older instruments were used, possibly hindering trends in our study by some amount.

>>>>

---
title: "Respiratory-Related Mortality Rates Show A Positive Correlation With Increasing Air Pollution"
subtitle: "Based on Data Collected From The Province Of Alberta"
author: 
  - Vanshika Vanshika
  - Shivank Goel
  - Navya Hooda
thanks: "Code and data are available at: https://github.com/shivankgoel003/Mortality-in-Alberta."
date: today
date-format: long
abstract: "Air pollution's severe impact on human respiratory and cardiac health is a growing public health concern across the world. The fine particulate matter (PM2.5) air pollutant is of particular concern because it can deeply penetrate the lungs and is largely found in urban environments. This research uses data from Alberta, Canada to assess any correlations between PM2.5 levels and the regional Air Quality Health Index (AQHI), against the number of deaths from groups of respiratory and cardiac illnesses we selected. We utilized statistical models and identified a strong association between increased PM2.5 levels and higher mortality from these diseases. These findings suggest that there is an immediate need for health policies that address air quality improvements to safeguard public health in Alberta and beyond."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

In our study, there might be several limitations affecting our conclusions.
First, between 2010 and 2017, older equipment at Alberta’s monitoring stations was replaced with new monitoring equipment for PM2.5. These new instruments measured an additional portion (semi-volatile) of the PM2.5 mass not captured by older instruments. As a result, the older monitoring equipment used between 2000 and 2017 likely underreported concentrations of PM2.5 under some conditions, after 2010, the increase in PM2.5 concentrations may be a result of changes in monitoring equipment @pm. Overall, the concentrations measured with the new monitors may not be directly comparable with measurements from years in which older instruments were used, possibly hindering trends in our study by some amount.
